### 2014-06-06 writing/S'14 paper/recursion.tex
More editing

In an attempt to better understand the limited success of the RNN model, we discovered that this model looses information about in longer statements in a particularly problematic way. It appears that that model is unable to distinguish between two sentences when the only difference between those sentences is embedded within a very deep structure. We evaluated both models on sentences that differed in only one term, but for which that one term was embedded under a large number of conjunctions, such as the pair \ii{a (and a)} and \ii{a (and b)}, or the pair \ii{a (and (a (and a)))} and \ii{a (and (a (and b)))}. We then measured the Euclidean distance between the vector representations of the two sentences in each pair. Our findings are shown in Figure~\ref{prop-falloff}, and show that while the RNTN can distinguish the two sentences well at every size that we test, the RNN fails after a depth of approximately six.

\begin{figure}[htp]
  \centering
  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
  \caption{How well each model can distinguish to expressions that differ by one term, measured for terms embedded under different numbers of tokens of \ii{and}.}  
  \label{prop-falloff}
\end{figure}

>>>>

In an effort to better understand why the RNTN outperforms the plain
RNN, we evaluated both models on pairs of long formulae involving
binary connectives, to assess the extent to which they achieve
substantially different representations for semantically distinct
formulae and substantially similar representations for synonymous
formulae. We found that both models are able to identify synonymous
pairs like $(a \pland a)$ and $(a \pland (a \pland (a \pland a)))$,
and similarly for disjunctions, but only the RNTN is able to learn
substantially different representations for formulae like $(a \pland
a)$ and $(a \pland (a \pland (a \pland
b)))$. Figure~\ref{prop-falloff} summarizes the result for pairs of
distinct formulae, showing that the RNN basically cannot distinguish
them after depth~2.

% we discovered that this model looses information about in longer
% statements in a particularly problematic way. It appears that that
% model is unable to distinguish between two sentences when the only
% difference between those sentences is embedded within a very deep
% structure. We evaluated both models on sentences that differed in only
% one term, but for which that one term was embedded under a large
% number of conjunctions, such as the pair \ii{a (and a)} and \ii{a (and
%   b)}, or the pair \ii{a (and (a (and a)))} and \ii{a (and (a (and
%   b)))}. We then measured the Euclidean distance between the vector
% representations of the two sentences in each pair. Our findings are
% shown in Figure~\ref{prop-falloff}, and show that while the RNTN can
% distinguish the two sentences well at every size that we test, the RNN
% fails after a depth of approximately six.

\begin{figure}[htp]
  \centering
  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
  \caption{Truth-conditionally distinct formulae should have different
    representations. As measured by Euclidean distance, only the RNTN
    achieves this for formulae containing more than two
    connectives. The RNN quickly collapses the representations,
    failing to capture the meaning contrasts.}
  \label{prop-falloff}
\end{figure}
